---
title: "Reproducible Research Week2 Project"
output: html_document
---

Let's load the packages.

```{r load-packages, message=FALSE}
library(ggplot2)
library(lattice)
```

## Load the data


```{r load, echo = TRUE}
activitydata<-read.csv("activity.csv",stringsAsFactors=FALSE)
# Transform the date attribute to an actual date format
activitydata$date <- as.POSIXct(activitydata$date, format="%Y-%m-%d")
```


### Create a new factor variables with two level-"weekday" and "weekend"
```{r weekday}
# Compute the weekdays from the date attribute
activity_1<- data.frame(date=activitydata$date, 
                           weekday=tolower(weekdays(activitydata$date)), 
                           steps=activitydata$steps, 
                           interval=activitydata$interval)

# Compute the day type (weekend or weekday)
activity_1 <- cbind(activity_1, 
                      daytype=ifelse(activity_1$weekday == "saturday" | 
                                     activity_1$weekday == "sunday", "weekend", 
                                     "weekday"))

# Create the final data.frame
activity <- data.frame(date=activity_1$date, 
                       weekday=activity_1$weekday, 
                       daytype=activity_1$daytype, 
                       interval=activity_1$interval,
                       steps=activity_1$steps)

# Clear the workspace
rm(activity_1)
```
## What is mean total number of steps taken per day?


```{r mean steps,echo = TRUE}
totalsteps<-tapply(activity$steps,activity$date, FUN = sum, na.rm = TRUE)
hist(totalsteps, col = "green", xlab = "Total number of steps taken each day")
```

```{r mean and median}
mean(totalsteps,na.rm = TRUE)
median(totalsteps, na.rm = TRUE)
```

## What is the avearge daily activity pattern?

```{r average}
average<-aggregate(x = list(meansteps = activity$steps), by = list(interval = activity$interval), FUN = mean, na.rm = TRUE)
ggplot(data = average, 
       aes(x = interval, y = meansteps))+ 
      geom_line()+ 
    xlab("5 min interval")+
    ylab("average number of steps taken")

```

```{r max}
average[which.max(average$meansteps),]
```


## Imputing missing values
### Calcuulate and report the missing values in dataset
```{r missing}
missing<-is.na(activitydata$steps)
table(missing)
```
### Devise a startegy for filling in all of the missing values and create a new dataset

We use the startegy of using the mean data for filling the missing values in datasets

```{r filling missing value}
napos<- which(is.na(activity$steps))
mean_vect<- rep(mean(activity$steps,na.rm = TRUE), times = length(napos))
activity[missing,"steps"]<- mean_vect
head(activity)
```



### Mke a histogram of the total number of steps taken and report mean and median per day.

```{r hist}
sum<-aggregate(activity$steps, by = list(interval = activity$date), FUN = sum, na.rm = TRUE)
names(sum)<- c("date","total")
hist(sum$total,
     breaks=seq(from=0, to=25000, by=2500),
     col = "green",
     xlab = "Total number of steps",
main = "Histogram of total number of steps")
```


```{r mean2 and median2}
mean(sum$total)
median(sum$total)
```


## Are there any difference in activity patterns between weekdays and weekdends?


### Make a panel plot containing a time series plot of 5 min interval and average nymber of steps taken

```{r another plot}
meandata<- aggregate(activity$steps, 
                       by=list(activity$daytype, 
                               activity$weekday, activity$interval), mean)

# Rename the attributes
names(meandata) <- c("daytype", "weekday", "interval", "mean")
xyplot(mean~interval|daytype,meandata,
       type = "l",
       xlab = "Interval",
       ylab = "Number of steps",
       layout = c(1,2))
```